---
title: "STAT UN 2103 Homework 3"
author: "Rebekah Kim"
date: "October 21, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Problem 1, 2 handwritten (back)

#### Problem 3

```{r}
estate_data <- read.table("HW3Problem3.txt", header = TRUE)
model <- lm(RentalRates~Age+OperatingExpense+VacancyRates+SquareFootage, data=estate_data)
model
```

i) Estimated linear model:

$Y = 12.2-0.142X_1+0.282X_2+0.6193X_3+7.924*10^{-6}X_4$

ii)
```{r}
summary(model)
```
There are marginal relationships between response variable and covariates (except Vacancy Rates), for their p-values are all < 0.05. Therefore we reject the null hypothesis for all covariates except Vacancy Rates, which has a high p-value > 0.05

iii)
```{r}
anova(model)
```
The p-values for f-test on rental rates and all of the covariates < 0.05
Therefore, we reject the null hypothesis. There is an overall relationsip between the response variable (Rental Rate) againt all covariates at 5% significance level.

#### Problem 4
```{r}
d <- read.table("HW3Problem4.txt", header = TRUE)

model <- lm(y~poly(x, 3, raw=TRUE), data=d)
model

summary(model)

plot(d$x,d$y, xlab="x", ylab="y", main="Estimated model for y~x")
curve(16.97286 + 88.16596*x-18.08742*x^2 +0.91547*x^3, add=TRUE)

```

estimated model:
$Y = 16.97286 + 88.16596X-18.08742 X^2 +0.91547X^3$


#### Problem 5
```{r}
# Data
y <- c(72,68,65,72,71,68,68,66,71,73)
x1 <- c(72,68,66,70,72,69,67,63,72,74)
x2 <- c(112,103,94,121,113,136,124,90,134,128)
# Sample size
n <- length(y)

# i) Design matrix
X <- c(rep(1,n),x1,x2)
dim(X) <- c(n,3)
X
```
$$\mathbf{X} = \left[\begin{array}
{rrr}
  1  & 72 & 112 \\
  1  & 68  &103\\
  1 &  66  & 94\\
   1 &  70  &121\\
  1 &  72  &113\\
   1  & 69 & 136\\
  1  & 67&  124\\
  1 &  63 &  90\\
  1   &72&  134\\
  1   &74 & 128
\end{array}\right]
$$

```{r}
# ii) Least squares 
beta <- solve(t(X)%*%X)%*%t(X)%*%y
beta

# iii)  SSE
H <- X%*%solve(t(X)%*%X)%*%t(X)
y.hat <- H%*%y
e <- y-y.hat

sse <- t(e)%*%e
sse

# iv) covariance matrix of beta
df <- n - 1
mse <- sse / df
sig.hatSq <- as.numeric(mse)

covbetamat <-sig.hatSq *solve(t(X)%*%X)
covbetamat

# v) filling out multiple regression summary tabl
model <- lm(y~x1+x2)
summary(model)

```


Estimate | Std.  Error | t value 
------------- | -------------|-------------
(Intercept) |17.762874  | 9.745144  | 1.823  
x1           |0.748239   |0.172550   |4.336  
x2          |-0.001869   |0.036288  |-0.051 




