---
title: "STAT UN2104 Project"
author: "Rebekah Kim"
date: "5/4/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

### Introduction
#### Research Quetion
> How is employee satisfaction related to available variables (religion, race, age, gender)?

We can construct a model to answer this question and predict which variables have statistically significant influence on employee satisfaction.

#### Set-up
```{r}
jobsat <- read.table("jobsat-data.txt", header=T)
jobsat["satprop"] <- jobsat$sat/(jobsat$sat + jobsat$unsat)

credit <- read.table("credit-data.txt", header=T)

```

#### Exploratory Data Analysis
```{r}
head(jobsat)
tail(jobsat)
str(jobsat)

ggplot(jobsat, aes(x= race, y = satprop)) +
  geom_boxplot() +
  facet_wrap(~age, nrow=1)


ggplot(jobsat, aes(x= gender, y = satprop)) +
  geom_boxplot() +
  facet_wrap(~region, nrow=1)
```

### Statistical Model
```{r}
sat.logit <- glm(formula=sat~factor(region)+race+factor(age)+gender, data=jobsat, family=binomial)
summary(sat.logit)
pchisq(q=6.8812-5.1116, df=10,lower.tail = F)
```

When all variables are entered, there are no significant 

### Executive Summary








