---
title: "STAT UN2103 HW4"
author: "Rebekah Kim"
date: "November 23, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Problem 1 (handwritten in back)

#### Problem 2

i)

$H_0 : \beta_1 = 0,\space H_A : \beta_3 \neq 0$
```{r}
d <- read.table("HW3Problem3.txt", header = TRUE)

model <- lm(RentalRates~Age+OperatingExpense+VacancyRates+SquareFootage, data = d)
summary(model)

```

$t_{calc} = 0.570, p-value = 0.57 > 0.05$

We fail to reject the null hypothesis; therefore, x3 (vacancy rate) is not statistically related to the response variable (rental rates) after holding all other variables constant.

ii)

$H_0 : \beta_1 = \beta_3 = 0, \space H_A : \beta_1 \neq 0 \space or \space \beta_3 \neq 0$

Full model:
$Y_i = \beta_0 + \beta_1x_{i1}+ \beta_2x_{i2}+ \beta_3x_{i3}+ \beta_4x_{i4}+ \epsilon_i$

$df_F = n - p = n - 5 = 81 - 5 = 76$

Reduced model:
$Y_i = \beta_0 + \beta_2x_{i2}+ \beta_4x_{i4}+ \epsilon_i$

$df_R = n - p = n - 3 = 81 - 3 = 78$

```{r}
r_model.1 <- lm(RentalRates~OperatingExpense+SquareFootage, data = d)
```
$f_{calc} = \dfrac{SSE_R - SSE_F}{df_R - df_F} \div \dfrac{SSE_F}{df_F}$

```{r}
#Shortcut
anova(r_model.1, model)
```

$SSE_F = 98.231$

$SSE_R = 159.491$

$f_{calc} = 23.968, \space p-value = 1.003e-08 < 0.05$

Therefore we reject the null hypothesis; age (x1) or vacancy rate (x3) may influence the rental rates.

iii)
$H_0 : \beta_3 = 0, \space H_A : \beta_3 \neq 0$

Full model:
$Y_i = \beta_0 + \beta_1x_{i1}+ \beta_2x_{i2}+ \beta_3x_{i3}+ \beta_4x_{i4}+ \epsilon_i$

$df_F = n - p = n - 5 = 81 - 5 = 76$

Reduced model:
$Y_i = \beta_0 + \beta_1x_{i1}+ \beta_2x_{i2}+ \beta_4x_{i4}+ \epsilon_i$

$df_R = n - p = n - 4 = 81 - 4 = 77$

```{r}
r_model.2 <- lm(RentalRates~Age+OperatingExpense+SquareFootage, data = d)
```
$f_{calc} = \dfrac{SSE_R - SSE_F}{df_R - df_F} \div \dfrac{SSE_F}{df_F}$

```{r}
#Shortcut
anova(r_model.2, model)
```

$SSE_F = 98.231$

$SSE_R = 98.650$

$f_{calc} = 0.3248, \space p-value = 0.5704 > 0.05$

Therefore, we fail to reject the null hypothesis; vacancy rate (x3) is not statistically related to the response variable (rental rates) after holding all other variables constant.
The f-calc (0.570) and p-value (0.57) here are the same as that of the t-test from part (i) and therefore we reach the same conclusion: fail to reject null hypothesis.

iv)


#### Problem 3


#### Problem 4


#### Problem 5 (handwritten in back)