---
title: "Applied Linear Regression Analaysis HW1"
author: "Rebekah Kim"
date: "September 20, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1


i. Are the regression assumptions satisfied in this application?

ii. Compute the sample correlation coe!cient of the above data set. Identify the strength
of the linear relationship.

iii. Compute the coe!cient of determination for this application. Interpret this quantity
relative to the problem statement.

iv. Find the line of best fit (LOBF) and the mean square error (MSE).

v. Interpret the estimated slope beta hat 1 in terms of the application.

vi. Predict the true average fungal growth when laetisaric acid concentration is fixed at 15
[µG/ml].

```{r}


```
## Problem 2

i. Use R to find the line of best fit for this application.
ii. Use R to find the correlation coe!cient for this application.
iii. Use R to create a scatter plot with the line of best fit. Make the line of best fit red.
iv. After looking at the scatter plot, do you think this line will do a good job predicting
alcohol consumption for a fixed level of tobacco usage?

## Problem 3
Consider a data set (x1, y1),(x2, y2),...,(xn, yn) and suppose yi = 2xi + 1 for i = 1, 2,...,n.
Show that the sample correlation coefficient r is equal to 1.

## Problem 4

i. Run an exploratory analysis on the above data set. This includes, finding descriptive
statistics of both data sets, construct histograms of the individual data sets, construct
a comparative box plot. Does the comparative box plot support the research question?
Make sure to add titles and label all axes on the plots.

```{r}

#datasets
control <- c(808,757,773,937,726,788,806,792,751,765,853,655,626,721,630,722,683,709,718,812,703,791,586,864,737,701,799,844,639,705,822,935,842,827,784,838,795,823,791,819) 
cell <- c(909,712,805,852,859,781,841,822,740,910,900,912,863,785,863,809,927,847,918,810,788,929,798,863,981,842,1021,827,876,736,640,851,787,703,942,758,843,781,969,872)

#Exploratory Analysis
#Descriptive Stat
summary(control)
mean(control)
var(control)

summary(cell)
mean(cell)
var(cell)
#Histogram
#Box plot


#Stacked data
y <- c(control,cell)

#Sample sizes
m <- length(control)
n <- length(cell)
```

ii. Are the assumptions satisfied to perform the relevant hypothesis test? To check this,
create QQ-plots of each data set and the stacked data set.

iii. Run a two-sample t-test on this data set. Include all relevant steps to receive full credit.
Note: You do not have to work out any formulas by hand. You can use R to find the
test statistic and p-value.

iv. Run a pooled two-sample t-test on this data set. Include all relevant steps to receive
full credit. Note: Again, you do not have to work out any formulas by hand. Also,
make sure to check that the variances do not statistically di↵er per group.

v. What type of error (Type I or Type II) did we maybe commit? Explain your reasoning
in one or two sentences.

vi. Should we use the pooled t-test or regular t-test for this application? Explain your
reasoning in one or two sentences.