---
title: "STAT UN2103 Section 001 Assignment #1"
author: "Rebekah Kim"
date: "September 20, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

i. The linear regression assumes 
  $Y_i = \beta_0 + \beta_1X_1 + \epsilon_i$, i = 1, 2, ... n
    
    where $\epsilon_i$ is independent and identifically distributed N(0, $\sigma^2$).
    
  Glancing at the fungus growth vs. laetisaric acid graph, there seems to be a linear relationship between the two variables; Indeed the variance of the error residuals ($e_i$) seems constant (the data points have similar lengths of deviation from the regression line). Also, the Q-Q plot further shows that it is plausible to assume that $\epsilon_i$ is normal with constant variance ($\epsilon_i \stackrel{iid}\sim N(0, \sigma^2)$).


ii.
$\hat{\beta}_1 = \dfrac{S_{xy}}{S_{xx}}$

$S_{xy} = \sum\limits_{i=1}^n (x_i-\bar{x})(y_i-\bar{y}) = -927.75$

$S_{xx} = \sum\limits_{i=1}^n (x_i-\bar{x})^2 = 1303$

$\hat{\beta}_1 = \dfrac{-927.75}{1303} = -0.712010744$

There is a relatively strong, negative (-0.71) linear relationship.


iii. coefficient of determination = $r^2 = 1 - \dfrac{SSE}{SST}$

$SSE = \sum\limits_{i=1}^n (y-\hat{y})^2 = 16.78$

$SST = S_{yy} = \sum\limits_{i=1}^n (y-\bar{y})^2 = 677.35$

$r^2 = 1 - \dfrac{SSE}{SST} = 1- 16.78/677.35 = 0.975226988$

$r^2$ is close to 0.98, meaning 98% of variation in fungus growth is explained by the laetisaric acid concentration. The acid compound probably effectively controls the fungus growth.


iv.

line of best fit ($\hat{y}$) = $\hat{\beta}_0 + \hat{\beta}_1x$

$\hat{\beta}_0 = \bar{y} - \hat{\beta}_1\bar{x}$ = 23.64 - (-0.712010744)*11.50 = 31.8281236

$\hat{y} = \hat{\beta}_0 + \hat{\beta}_1x$ =  31.8281236 -0.712010744x

MSE = $\dfrac{SSE}{n-2} = \dfrac{16.78}{12-2}$ = 1.678



v.

For every 1 $\mu$ G/mL increase in laetisaric acid concentration, the average fungus growth (radial measurement of the colony) decrease by 0.7120107mm



vi. Predict the true average fungal growth when laetisaric acid concentration is fixed at 15
[µG/ml].

31.8281236 -0.712010744*15 = 21.14796 mm of fungal growth

## Problem 2
i. Use R to find the line of best fit for this application.
```{r}
#dataset
alcohol <- c(6.47, 6.13, 6.19, 4.89, 5.63, 4.52, 5.89, 4.79, 5.27, 6.08, 4.02)
tobacco <- c(4.03, 3.76, 3.77, 3.34, 3.47, 2.92, 3.20, 2.71, 3.53, 4.51, 4.56)

#linear model
model = lm(alcohol~tobacco)

#find beta 1 (slope) and beta 0 (intercept) of line of best fit
model
```
line of best fit ($\hat{y}$) = $\hat{\beta}_0 + \hat{\beta}_1x$
  = 4.3512 + 0.3019x


ii. Use R to find the correlation coefficient for this application.
```{r}
cor(tobacco, alcohol)
```

iii. Use R to create a scatter plot with the line of best fit. Make the line of best fit red.
```{r}
plot(tobacco, alcohol, main="Alcohol vs Tobacco expenditure")
abline(model, col=2)
```

iv. After looking at the scatter plot, do you think this line will do a good job predicting
alcohol consumption for a fixed level of tobacco usage?
```{r}
summary(model)

summary(model)$r.squared
```

No this line will probably not do a good job because there is weak correlation (0.2236), low R-squared value (0.05), and a high p-value (0.5087).

## Problem 3
Consider a data set (x1, y1),(x2, y2),...,(xn, yn) and suppose yi = 2xi + 1 for i = 1, 2,...,n.
Show that the sample correlation coefficient r is equal to 1.

```{r}
x <- 1:1000
y <- 2 * x + 1
cor(x, y)
```

## Problem 4

i. Run an exploratory analysis on the above data set. This includes, finding descriptive
statistics of both data sets, construct histograms of the individual data sets, construct
a comparative box plot. Does the comparative box plot support the research question?
Make sure to add titles and label all axes on the plots.

```{r}

#datasets
control <- c(808,757,773,937,726,788,806,792,751,765,853,655,626,721,630,722,683,709,718,812,703,791,586,864,737,701,799,844,639,705,822,935,842,827,784,838,795,823,791,819) 
cell <- c(909,712,805,852,859,781,841,822,740,910,900,912,863,785,863,809,927,847,918,810,788,929,798,863,981,842,1021,827,876,736,640,851,787,703,942,758,843,781,969,872)

#Exploratory Analysis
#Descriptive Statistics on control and cellphone datasets
summary(control)
mean(control)
var(control)

summary(cell)
mean(cell)
var(cell)

#Histogram
hist(control)
hist(cell)

#Stacked data
y <- c(control,cell)

#Box plot
boxplot(y~c(rep("control", 40), rep("cell", 40)), main="Box plot", ylab = "reaction time")
boxplot (control, cell, main="Box plot", ylab = "Reaction time")



#Sample sizes
m <- length(control)
n <- length(cell)
```

ii. Are the assumptions satisfied to perform the relevant hypothesis test? To check this,
create QQ-plots of each data set and the stacked data set.

iii. Run a two-sample t-test on this data set. Include all relevant steps to receive full credit.
Note: You do not have to work out any formulas by hand. You can use R to find the
test statistic and p-value.

iv. Run a pooled two-sample t-test on this data set. Include all relevant steps to receive
full credit. Note: Again, you do not have to work out any formulas by hand. Also,
make sure to check that the variances do not statistically di↵er per group.

v. What type of error (Type I or Type II) did we maybe commit? Explain your reasoning
in one or two sentences.

vi. Should we use the pooled t-test or regular t-test for this application? Explain your
reasoning in one or two sentences.